### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@3bee2b6e9b1d6cc01c9e2dc9e73ad7bf73394406 🚀

Spillover risk is multidimensional and complicated. From the standpoint
of an animal community, \emph{i.e.} a pool of suitable hosts, it is
driven by a multiplicity of factors (Plowright et al. 2017). Proxies for
the global richness of hosts is one such component commonly analysed
(see \emph{e.g.} Anthony et al. 2017 for coronaviruses), but there is an
argument to be made that species who are not competent (or known) hosts
of a specific virus genus may not factor into this (Plowright et al.
2015); this is especially true as competence data increases predictive
power when the taxonomic scope of hosts of a viral family increases
(\textbf{Becker2020InfInt?}, \textbf{Mull2022VirIso?}). Similarly, host
species who are assumed to share viruses at different rates should be
weighted accordingly (Albery et al. 2020). In mammals, key functional
traits (for which phylogeny is a reasonable proxy) are determinants of
the spillover potential (Olival et al. 2017); these include, notably,
body mass and affinity for urban environments (Albery et al. 2022).
Finally, especially when the pool of potential hosts spans the entire
globe, there may be local host pools that are highly unique; not having
been observed in other locations, these can act on the overall risk
either by providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid evolutionary changes in specialism of their pathogens (Agosta et
al. 2010). In the specific case of generalist pathogens (which is the
case many viruses in the betacoronavirus genus, see \emph{e.g.}
\textbf{MacLean2021NatSel?}), there is conceptual and empirical support
to the idea that these community-level mechanisms are even more
important in driving the overall risk (Power and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
threaten human health (\textbf{Letko2020BatVir?},
\textbf{VanBrussel2022ZooDis?}). Chiropterans emerged around 64 million
years ago and are one of the most diverse mammalian orders, with an
estimated richness of more than 1400 species (Peixoto et al. 2018,
Simmons and Cirranello 2020). They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to human (Kasso and Balakrishnan
2013). For example, bats are an essential component of many
seed-dispersal networks (\textbf{Mello2011MisPar?}). Over two-thirds of
bats are know to be either obligate or facultative insectivores,
therefore playing an important role in the regulation of insect pests
that can affect crops (Williams-Guillén et al. 2008, Voigt and Kingston
2016), and vectors of pathogens that put a risk on human health
(Gonsalves et al. 2013a, b). Because bats are globally distributed and
have a long evolutionary history, phylogeographic and biogeographic
approaches are required to shed light on the contemporary distribution
of coevolutionary processes between bats and the pathogens they host.
Not all areas in which bats, viruses, and human are co-occuring are
facing a risk of spillover towards human populations, and the areas in
which this risk exist may not be facing risks of the same nature and
magnitude.

@@

Bats are found worldwide and are both one of the most diverse groups
among mammals (\textbf{Moratelli \& Calisher, 2015}), and one of the
main animal reservoir for different strains of betacoronaviruses
(Drexler et al. 2014). This has attracted attention to areas where high
diversity of bats, and therefore presumably high diversity of
betacoronaviruses, can be an important issue for human health (Calisher
et al. 2006, Moratelli and Calisher 2015). By overlaying the IUCN
rangempas for confirmed bat hosts of betacoronaviruses
{[}fig.~\ref{fig:richness}; top{]}, we see that the the main hotspots of
host richness are primarily South-Eastern Asia, parts of Southern
Europe, and to a lesser extent parts of Africa in the -25-0 range of
latitudes. The description of host richness is an important first step
towards understanding risk, as previous research (\textbf{Anthony et
al., 2017; Mollentze \& Streicker, 2020}) states that locally diverse
bat communities could maintain more viruses and hence, a higher
probability of having a pathogen that could represent a risk for human
health.

@@

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronavirus
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data; furthermore, there are more sources of information on hosts
than on viruses, allowing to develop a host-centric perspective on risk.
Areas with high bat diversity and high turnover \emph{may} facilitate
the evolutionary radiation of viruses, matching previous findings that
the diversification of bat coronaviruses is driven largely by host
shifts (inter-genus or higher levels of cross-species transmission) and,
to a lesser degree, cospeciation and sharing (intra-genus cross-species
transmission; \textbf{Anthony et al.~2017}). This diversification is not
an actual risk factor for spillover itself, but acts downstream of a
spillover event by increasing the random chance of the emergence of a
virus with the raw genomic components required for the potential to
infect humans.

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

Spillover risk is multidimensional and complicated. From the standpoint
of an animal community, \emph{i.e.} a pool of suitable hosts, it is
driven by a multiplicity of factors (Plowright et al. 2017). Proxies for
the global richness of hosts is one such component commonly analysed
(see \emph{e.g.} Anthony et al. 2017 for coronaviruses), but there is an
argument to be made that species who are not competent (or known) hosts
of a specific virus genus may not factor into this (Plowright et al.
2015); this is especially true as competence data increases predictive
power when the taxonomic scope of hosts of a viral family increases
(Becker et al. 2020, Mull et al. 2022). Similarly, host species who are
assumed to share viruses at different rates should be weighted
accordingly (Albery et al. 2020). In mammals, key functional traits (for
which phylogeny is a reasonable proxy) are determinants of the spillover
potential (Olival et al. 2017); these include, notably, body mass and
affinity for urban environments (Albery et al. 2022). Finally,
especially when the pool of potential hosts spans the entire globe,
there may be local host pools that are highly unique; not having been
observed in other locations, these can act on the overall risk either by
providing novel contact opportunities, reflecting unique
host-environment combinations (Engering et al. 2013), or facilitating
rapid evolutionary changes in specialism of their pathogens (Agosta et
al. 2010). In the specific case of generalist pathogens (which is the
case many viruses in the betacoronavirus genus, see \emph{e.g.} MacLean
et al. 2021), there is conceptual and empirical support to the idea that
these community-level mechanisms are even more important in driving the
overall risk (Power and Mitchell 2004).

Bats are important reservoir hosts for different classes of
microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
threaten human health (Letko et al. 2020, Van Brussel and Holmes 2022).
Chiropterans emerged around 64 million years ago and are one of the most
diverse mammalian orders, with an estimated richness of more than 1400
species (Peixoto et al. 2018, Simmons and Cirranello 2020). They exhibit
a broad variety of habitat use, behaviour, and feeding strategies,
putting them at key positions in the delivery and provisioning of
several ecosystem services, tied to important ecosystem-derived benefits
to human (Kasso and Balakrishnan 2013). For example, bats are an
essential component of many seed-dispersal networks (Mello et al. 2011).
Over two-thirds of bats are know to be either obligate or facultative
insectivores, therefore playing an important role in the regulation of
insect pests that can affect crops (Williams-Guillén et al. 2008, Voigt
and Kingston 2016), and vectors of pathogens that put a risk on human
health (Gonsalves et al. 2013a, b). Because bats are globally
distributed and have a long evolutionary history, phylogeographic and
biogeographic approaches are required to shed light on the contemporary
distribution of coevolutionary processes between bats and the pathogens
they host. Not all areas in which bats, viruses, and human are
co-occuring are facing a risk of spillover towards human populations,
and the areas in which this risk exist may not be facing risks of the
same nature and magnitude.

@@

Bats are found worldwide and are both one of the most diverse groups
among mammals (\textbf{Moratelli \& Calisher, 2015}), and one of the
main animal reservoir for different strains of betacoronaviruses
(Drexler et al. 2014). This has attracted attention to areas where high
diversity of bats, and therefore presumably high diversity of
betacoronaviruses, can be an important issue for human health (Calisher
et al. 2006, Moratelli and Calisher 2015). By overlaying the IUCN
rangempas for confirmed bat hosts of betacoronaviruses
{[}fig.~\ref{fig:richness}; top{]}, we see that the the main hotspots of
host richness are primarily South-Eastern Asia, parts of Southern
Europe, and to a lesser extent parts of Africa in the -25-0 range of
latitudes. The description of host richness is an important first step
towards understanding risk, as previous research (Anthony et al. 2017,
Mollentze and Streicker 2020) states that locally diverse bat
communities could maintain more viruses and hence, a higher probability
of having a pathogen that could represent a risk for human health.

@@

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronavirus
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In
fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
areas, which are essentially limited to South-Eastern Asia, and to some
part of Sub-Saharan Africa. Adopting a geographic mosaic theory
perspective on risk, other regions of the world are of lesser concern
fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
diversity or distinctiveness. The simple rationale behind it is that the
acquisition of viral data is rarely disconnected from the acquisition of
host data; furthermore, there are more sources of information on hosts
than on viruses, allowing to develop a host-centric perspective on risk.
Areas with high bat diversity and high turnover \emph{may} facilitate
the evolutionary radiation of viruses, matching previous findings that
the diversification of bat coronaviruses is driven largely by host
shifts (inter-genus or higher levels of cross-species transmission) and,
to a lesser degree, cospeciation and sharing (intra-genus cross-species
transmission; Anthony et al. 2017). This diversification is not an
actual risk factor for spillover itself, but acts downstream of a
spillover event by increasing the random chance of the emergence of a
virus with the raw genomic components required for the potential to
infect humans.

@@

From another perspective, areas of high host uniqueness and virus
sharing (red-to-pink) could provide hotspots of betacoronavirus risk
through mixing of unique viruses (via codivergence) and in turn
recombination. Under our framework, such a hotspot was identified in
Madagascar, where most bat species are endemic following evolutionary
divergence from sister species in both African and Asian continents
(\emph{e.g.} Shi et al. 2014). Recent surveillance (Kettenburg et al.
2022) has identified a novel betacoronavirus (in the subgenus
\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
strong proof of principle in model predictions.

@@

\leavevmode\hypertarget{ref-Becker2020InfInt}{}%
Becker, D. J. et al. 2020. Beyond Infection: Integrating Competence into
Reservoir Host Prediction. - Trends in Ecology \& Evolution 35:
1062--1065.

@@

\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
Kettenburg, G. et al. 2022. Full Genome Nobecovirus Sequences From
Malagasy Fruit Bats Define a Unique Evolutionary History for This
Coronavirus Clade. - Frontiers in Public Health in press.

@@

\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
Letko, M. et al. 2020. Bat-borne virus diversity, spillover and
emergence. - Nature Reviews Microbiology 18: 461--471.

\leavevmode\hypertarget{ref-MacLean2021NatSel}{}%
MacLean, O. A. et al. 2021. Natural selection in the evolution of
SARS-CoV-2 in bats created a generalist virus and highly capable human
pathogen. - PLOS Biology 19: e3001115.

@@

\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
Mello, M. A. R. et al. 2011. The Missing Part of Seed Dispersal
Networks: Structure and Robustness of Bat-Fruit Interactions. - PLOS ONE
6: e17395.

\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
Mollentze, N. and Streicker, D. G. 2020. Viral zoonotic risk is
homogenous among taxonomic orders of mammalian and avian reservoir
hosts. - Proceedings of the National Academy of Sciences 117:
9423--9430.

@@

\leavevmode\hypertarget{ref-Mull2022VirIso}{}%
Mull, N. et al. 2022. Virus isolation data improve host predictions for
New World rodent orthohantaviruses. - Journal of Animal Ecology in
press.

@@

\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
Shi, J. J. et al. 2014. A Deep Divergence Time between Sister Species of
Eidolon (Pteropodidae) with Evidence for Widespread Panmixia. - Acta
Chiropterologica 16: 279--292.

@@

\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
Van Brussel, K. and Holmes, E. C. 2022. Zoonotic disease and virome
diversity in bats. - Current Opinion in Virology 52: 192--202.
